---
title: "Signature Comparison"
date: "12/16/2024"
layout: full
format:
  html: default
toc: true
toc-location: right
embed-resources: TRUE
---

This code relies on the cytoverse packages for getting the raw .fcs files into GatingSet objects for subsequent handling. These are then passed to Luciernaga (currently available via GitHub) for the signature processing and plotting normalized signatures.  

# Setup

## Load Libraries
```{r}
#| code-fold: show
#| warning: FALSE
#| message: false
# CRAN packages (install.packages("ThisOne"))
library(data.table)
library(gt)
library(stringr)
library(dplyr)

# Bioconductor packages( BiocManager::install("ThisOne"))
library(flowCore)
library(flowWorkspace)
library(openCyto)
library(ggcyto)
```

```{r}
#| eval: FALSE
remotes::install_github("https://github.com/DavidRach/Luciernaga")
library(Luciernaga)
```

## Locating FCS Files
Using .fcs files from within the Luciernaga R package for this example, for your case specify the location of the raw fcs file folder by swapping in the #ed argument. 

```{r}
#| code-fold: show
File_Location <- system.file("extdata", package = "Luciernaga")
#Alternately: File_Location <- file.path("C:", "Users", "JohnDoe", "Desktop", "TodaysExperiment")
FCS_Pattern <- ".fcs$"
FCS_Files <- list.files(path = File_Location, pattern = FCS_Pattern,
                        full.names = TRUE, recursive = FALSE)
```

```{r}
#| code-fold: show
UnstainedFCSFiles <- FCS_Files[grep("Unstained", FCS_Files)]
UnstainedBeads <- UnstainedFCSFiles[grep("Beads", UnstainedFCSFiles)]
UnstainedCells <- UnstainedFCSFiles[-grep("Beads", UnstainedFCSFiles)]

BeadFCSFiles <- FCS_Files[grep("Beads", FCS_Files)]
BeadSingleColors <- BeadFCSFiles[-grep("Unstained", BeadFCSFiles)]

CellSingleColors <- FCS_Files[grep("Cells", FCS_Files)]
```

## Gating For Single-Colors Cells
```{r}
#| code-fold: show
MyCytoSet <- load_cytoset_from_fcs(CellSingleColors, truncate_max_range = FALSE,
                                   transformation = FALSE)
MyCellGatingSet <- GatingSet(MyCytoSet)
MyCellGatingSet
```

```{r}
#| code-fold: show
FileLocation <- system.file("extdata", package = "Luciernaga")
# FileLocation <- file.path("SignatureComparison")
MyGates <- fread(file.path(path = FileLocation, pattern = 'Gates.csv'))
gt(MyGates)
```

```{r, message=FALSE}
#| code-fold: show
#| message: false
MyGatingTemplate <- gatingTemplate(MyGates)
gt_gating(MyGatingTemplate, MyCellGatingSet)

#pData(MyCellGatingSet[[20]])

BV711_Cell <- subset(MyCellGatingSet, str_detect(name, "BV711"))
BV711_Cell[[1]]

autoplot(BV711_Cell, "lymphocytes")
```

```{r}
#| eval: FALSE
#| code-fold: show
# Alternately visualize gate placement via Luciernaga

if(!require("remotes")) install.packages("remotes")
remotes::install_github("https://github.com/DavidRach/Luciernaga")

library(Luciernaga)

removestrings <-  c("DR_", "Cells", ".fcs", "-", " ")
StorageLocation <- file.path("C:", "Users", "JohnDoe", "Desktop")

IndividualPlot <- Utility_GatingPlots(x=BV711_Cell[[1]], sample.name = "GUID",
                                      removestrings = removestrings,
                                      gtFile = MyGates, DesiredGates = NULL,
                                      outpath = StorageLocation, export = FALSE,
                                      thecolumns=2, therows=3)
IndividualPlot[[1]]
```

## Gating For Single-Color Beads

```{r}
#| code-fold: show
MyBeadsCytoSet <- load_cytoset_from_fcs(BeadSingleColors,
                                        truncate_max_range = FALSE,
                                        transform = FALSE)

MyBeadsGatingSet <- GatingSet(MyBeadsCytoSet)
```

```{r, message=FALSE}
#| code-fold: show
FileLocation <- system.file("extdata", package = "Luciernaga")
# FileLocation <- file.path("SignatureComparison")
MyBeadsGates <- fread(file.path(path = FileLocation, 
                                pattern = 'GatesBeads.csv'))
gt(MyBeadsGates)
```

```{r}
#| message: FALSE
#| code-fold: show
MyBeadsGatingTemplate <- gatingTemplate(MyBeadsGates)
gt_gating(MyBeadsGatingTemplate, MyBeadsGatingSet)

BV711_Bead <- subset(MyBeadsGatingSet, str_detect(name, "BV711"))
BV711_Bead[[1]]
plot(BV711_Bead)

autoplot(BV711_Bead, "beads")
```

```{r}
#| eval: FALSE
#| code-fold: show
# Alternately visualize gate placement via Luciernaga

# if(!require("remotes")) install.packages("remotes")
# remotes::install_github("https://github.com/DavidRach/Luciernaga")

library(Luciernaga)

removestrings <-  c("DR_", "Cells", ".fcs", "-", " ")
StorageLocation <- file.path("C:", "Users", "JohnDoe", "Desktop")

IndividualPlot <- Utility_GatingPlots(x=BV711_Bead[[1]], sample.name = "GUID",
                                      removestrings = removestrings,
                                      gtFile = MyBeadsGates, DesiredGates = NULL,
                                      outpath = StorageLocation, export = FALSE,
                                      thecolumns=2, therows=3)
IndividualPlot[[1]]
```

## Gating for Unstained Cells

```{r}
#| code-fold: show
# UnstainedCells # Selecting first specimen in list
MyUnstainedCytoSet <- load_cytoset_from_fcs(UnstainedCells[1], 
                                   truncate_max_range = FALSE, 
                                   transform = FALSE)
MyUnstainedGatingSet <- GatingSet(MyUnstainedCytoSet)
```

```{r}
#| code-fold: show
FileLocation <- system.file("extdata", package = "Luciernaga")
# FileLocation <- file.path("SignatureComparison")
MyGates <- fread(file.path(path = FileLocation, pattern = 'Gates.csv'))
gt(MyGates)
```

```{r}
#| code-fold: show
#| message: false
MyGatingTemplate <- gatingTemplate(MyGates)
gt_gating(MyGatingTemplate, MyUnstainedGatingSet)
MyUnstainedGatingSet[[1]]
autoplot(MyUnstainedGatingSet[[1]], "lymphocytes")
```

## Gating for Unstained Beads

```{r}
#| code-fold: show
MyUnstainedBeadsCytoSet <- load_cytoset_from_fcs(UnstainedBeads,
                                        truncate_max_range = FALSE,
                                        transform = FALSE)
MyUnstainedBeadsGatingSet <- GatingSet(MyUnstainedBeadsCytoSet)
```

```{r}
#| code-fold: show
FileLocation <- system.file("extdata", package = "Luciernaga")
# FileLocation <- file.path("SignatureComparison")
MyBeadsGates <- fread(file.path(path = FileLocation, 
                                pattern = 'GatesBeads.csv'))
gt(MyBeadsGates)
```

```{r}
#| code-fold: show
#| message: false
MyBeadsGatingTemplate <- gatingTemplate(MyBeadsGates)
gt_gating(MyBeadsGatingTemplate, MyUnstainedBeadsGatingSet)
MyUnstainedBeadsGatingSet[[1]]
plot(MyUnstainedBeadsGatingSet)
pData(MyUnstainedBeadsGatingSet)
autoplot(MyUnstainedBeadsGatingSet, "beads")
```

# Extracting Data

```{r}
#| code-fold: show
#| message: false
library(Luciernaga)

FileLocation <- system.file("extdata", package = "Luciernaga")
# FileLocation <- file.path("SignatureComparison")
pattern = "AutofluorescentOverlaps.csv"
AFOverlap <- list.files(path=FileLocation, pattern=pattern,
                        full.names = TRUE)
removestrings <- c(".fcs")
StorageLocation <- file.path("C:", "Users", "JohnDoe", "Desktop")

BV711Cell_Signatures  <- Luciernaga_QC(x=BV711_Cell[1],
                                      subsets="lymphocytes",
                                      removestrings=removestrings,
                                      sample.name="GUID",
                                      unmixingcontroltype = "cells",
                                      Unstained = FALSE,
                                      ratiopopcutoff = 0.01,
                                      Verbose = TRUE,
                                      AFOverlap = AFOverlap, 
                                      stats = "median",
                                      ExportType = "data",
                                      SignatureReturnNow = FALSE,
                                      outpath = NULL, Increments=0.1,
                                      experiment="Test", condition="Cells")

UnstainedBeads <- Luciernaga_QC(x=MyUnstainedBeadsGatingSet[1], desiredAF = NULL,
                    subsets="singlets", removestrings=removestrings,
                    sample.name="GUID", unmixingcontroltype = "beads",
                    Unstained = TRUE, ratiopopcutoff = 0.001, Verbose = TRUE,
                    AFOverlap = AFOverlap, stats = "median",
                    ExportType = "data.frame", SignatureReturnNow = TRUE,
                    outpath = NULL, experiment="Test", condition="Test")

BV711Bead_Signatures <- Luciernaga_QC(BV711_Bead[1], subsets="singlets",
                              removestrings=removestrings, sample.name="GUID",
                              unmixingcontroltype = "cells", Unstained = FALSE,
                              ratiopopcutoff = 0.001, Verbose = FALSE,
                              AFOverlap = AFOverlap, stats = "median",
                              ExportType = "data", SignatureReturnNow = FALSE,
                              outpath = StorageLocation, Increments=0.1,
                              SecondaryPeaks=2, experiment = "Test",
                              condition = "Beads", SCData="subtracted",
                              NegativeType="default", BeadAF=UnstainedBeads,
                              BeadMainAF="UV1-A")
```

```{r}
#| code-fold: show

# Minor Data Wrangling Due to A Bug I Am Dealing With

Data <- rbind(BV711Cell_Signatures, BV711Bead_Signatures) %>% mutate(Signature=paste(Sample, Cluster)) %>% relocate(Signature, .before=1) %>% select(-Sample, -Experiment, -Condition, -Cluster, -Count)
colnames(Data)[1] <- "Sample"
TheSignatures <- Data %>% pull(Sample)

Plot <- Luciernaga::QC_ViewSignature(x=TheSignatures, data=Data, Normalize=TRUE)

plotly::ggplotly(Plot)
```

